### 2017-01-09 introduction.tex
Initial Commit

\section{Introduction}

This file contains comments made to papers. The comments are structured as five
paragraph essays about those papers.

Each paper is identified by the code \texttt{IND\_XXX}, where \texttt{XXX} is a
number with 3 leading zeros. Each paper is also identified by a reference in
the references section. As some papers reference each other, these references
are kept in the comments.