### 2017-02-10 exps.tex

\chapter{Experiments}

\section{EXP\_001 - Plotting Activity of Students over the Semester}

\subsection{Objective}

To look for patterns in the usage of Moodle we look into plotting the activity
of students over the courses over a semester. There are many ways of doing so.

One idea is to plot the average number of activities of students on one course.
Over a semester, several students access a specific course where they read and
write forum messages, deliver assignments, etc. For the various courses, we
plot the average of the total number of activities, and the average number of
each individual~\gls{crud} activities (\textbf{C}reate, \textbf{R}ead, etc).

\subsection{Procedure}

\subsection{Conclusions}

Some graphs show distinct peaks. What are they fore? Is it for a new
assignment, or just before an evaluation?

Some courses have a greater number of peaks them others. Do these courses have
more students, more assignments, better grades?