### 2017-02-24 ResearchNotes/introduction.tex
Updated Research Notes

\chapter{Introduction}

The objective of this project is to mine Moodle's data to find if user
activities explains their success. The hypothesis is that some behaviours in
courses during a semester may lead to success or lack there off.

This document contains the research notes made during the development of this
project. It contains a section on background, which has comments done to
papers, a listing of relevant journals and books, and some state of the art
regarding Data Mining, Educational Data Mining, and Moodle Mining.

The document also contains a section dedicate to the data which is used in this
project and a section detailing statistical analysis and experiments done to
that data.