### 2017-02-28 ResearchNotes/books.tex
Added info on books

\section{Books}

Here is a listing of books which were relevant for this project. Each section
describes a single book, and how that particular book was relevant in this
project.

\subsection{``Machine Learning''}

This book is~\cite{book_ml} and is titled ``Machine Learning''. Although this
is a Data Mining project, many of the algorithms and methods exposed by this
book are very relevant in Data Mining, namely chapter 3 which is on Decision
Trees and chapter 8 which is on Instance Based Learning.

\subsection{``Data Mining and Knowledge Discovery Handbook''}

Book is~\cite{book_dm_kdh}.~\TODO{}

\subsection{``Introduction to Data Mining''}

Book is~\cite{book_intro_dm}.~\TODO{}

\subsection{``Data Mining: Practical Machine Learning Tools and Techniques''}

Book is~\cite{book_dm_practical}.~\TODO{}