### 2017-03-01 ResearchNotes/books.tex
Adds papers and books descriptions to notes

\section{Books}

Here is a listing of books which were relevant for this project. Each section
describes a single book, and how that particular book was relevant in this
project.

\subsection{``Machine Learning''}

This book is~\cite{book_ml} and is titled ``Machine Learning''. Although this
is a Data Mining project, many of the algorithms and methods exposed by this
book are very relevant in Data Mining, namely chapter 3 which is on Decision
Trees and chapter 8 which is on Instance Based Learning.

\subsection{``Data Mining and Knowledge Discovery Handbook''}

Book is~\cite{book_dm_kdh}.~\TODO{}

\subsection{``Introduction to Data Mining''}

Book is~\cite{book_intro_dm}.~\TODO{}

\subsection{``Data Mining: Practical Machine Learning Tools and Techniques''}

This Book~\cite{book_dm_practical} is very practical not only because it talk
about many models, but also because is talk about how Machine Learning and data
Mining relate. Such notion might be important in the thesis text or, at the
very least, it might be important is someone asks why I cite sources about
Machine Learning.

The usual classifier models, such as decision trees, and usual clustering
methods are specified in this book.