### 2017-03-15 ResearchNotes/introduction.tex

\chapter{Introduction}

This document contains the research notes made during the development of the
thesis called ``Educational Data Mining applied to Moodle data from the
University of Évora''.

Initially, there are a few comments made to papers and books. Some journals are
also mentioned.

The document has a detailed explanation about the original data and how it was
processed into a data repository for experiments.

Information about statistics, processed datasets, and experiments are in their
own section.

Note that this document doesn't have all the information for the thesis. Some
text in the thesis was copy-pasted from this initial document, but a lot of
details are only added or corrected in the thesis text.