### 2017-03-23 ResearchNotes/tools.tex

\section{Tools}

Python 3.5.1 was used for everything. The tools used were:

\begin{itemize}
    \item Numpy 1.11.3~\cite{web_numpy}
    \item SciPy 0.18.1~\cite{web_scipy}
    \item SciKit-Learn 0.18.1~\cite{web_scikit}
    \item Pandas 0.19.2~\cite{web_pandas}
    \item Matplotlib 1.5.4~\cite{web_matplotlib}
\end{itemize}