### 2017-03-23 ResearchNotes/tools.tex
Adds information on tools

\section{Tools}

Python 3.5.1 was used for everything. The tools used were:

\begin{itemize}
    \item Numpy 1.11.3~\cite{web_numpy}
    \item SciPy 0.18.1~\cite{web_scipy}
    \item SciKit-Learn 0.18.1~\cite{web_scikit}
    \item Pandas 0.19.2~\cite{web_pandas}
    \item Matplotlib 1.5.4~\cite{web_matplotlib}
\end{itemize}

Documentation pages:

\begin{itemize}
    \item Decision Trees in SciKit-Learn~\cite{web_scikit_trees}
    \item Bayesian Networks in SciKit-Learn~\cite{web_scikit_bayes}
    \item K-Means in SciKit-Learn~\cite{web_scikit_kmeans}
    \item Affinity Propagation in SciKit-Learn~\cite{web_scikit_affinity}
\end{itemize}