### 2017-03-28 ThesisText/final.tex
Corrects small details

\chapter{Conclusions and Future Work}
\label{sec:final}

% =============================================================================
\section{This Project}
\TODO{}

In this project~\gls{edm} techniques were applied to the University of Évora's
Moodle data. The tasks applied to the data were based on similar projects shown
in the background section which also applied~\gls{edm} techniques to Moodle
data of other learning institutions.

In this project it was shown how the original data from the University of
Évora's Moodle was preprocessed into a data repository useful for analysis.
Some statistics over that data were then shown.

The techniques applied are divided into Supervised and Unsupervised learning. In supervised learning

% =============================================================================
\section{Future Work}
\TODO{}