### 2017-07-05 presentation/final.tex

\section{Conclusions}

% =============================================================================
\begin{frame}
\frametitle{Conclusions}

\center{\textbf{Recap}}

\vspace{0.5cm}

\begin{itemize}
    \item Objects and their features were extracted from the data
    \item Statistics were calculated
    \item Models were trained using that data
\end{itemize}

\center{\textbf{Results}}

\vspace{0.5cm}

\begin{itemize}
    \item It is possible to make certain prediction about student success based on Moodle Data
    \item Some clusters exits in the data
\end{itemize}

\end{frame}

% =============================================================================
\begin{frame}
\frametitle{Conclusions}

\center{\textbf{Future Work}}

\begin{itemize}
    \item Further analyses to clustering results
    \item Train other models such as Association Rules and Neural Networks
    \item Find why some instances of training have very low accuracy values
\end{itemize}

\vspace{0.5cm}

\center{\textbf{Applications}}

\begin{itemize}
    \item Make systems in which the success of students is predicted before he finishes a course
    \item Such system may signal professors and students about the possibility of failure of some people in courses
\end{itemize}

\end{frame}

% =============================================================================
\begin{frame}
\frametitle{Conclusions}

\center{\textbf{Thank you for Listening}}

\end{frame}