### 2017-07-05 presentation/intro.tex

\section{Introduction}

% =============================================================================
\begin{frame}
\frametitle{Introduction}

\center{\textbf{Project Consists On}}

\vspace{0.5cm}

\begin{itemize}
    \item Exploring data from the Moodle of the University of Évora
    \item Applying Educational Data Mining (EDM) methods to that data
\end{itemize}

\center{\textbf{Presented Over the Following Steps}}

\vspace{0.5cm}

\begin{itemize}
    \item Looking into what E-Learning is and how it is implemented
    \item Exploring Data Mining and Educational Data Mining
    \item Analysing the data from Évora's Moodle
    \item Training models to find interesting patterns in data
\end{itemize}

\end{frame}

% =============================================================================
\begin{frame}
\frametitle{Background}

\center{\textbf{E-Learning, LMSs, and Moodle}}

\begin{itemize}
    \item E-Learning is a way of delivering education and training remotely
    \item Due to advancements in technologies, E-Learning has been becoming more popular
    \item Learning Management Systems (LMS), like Moodle, have been successfully used to implement E-Learning
    \item These systems generate lots of data which are not trivially analysed
\end{itemize}

\end{frame}

% =============================================================================
\begin{frame}
\frametitle{Background}

\center{\textbf{Data Mining}}

\begin{itemize}
    \item DM aims at finding patterns in data in order to make non trivial assumptions
    \item This is done by training models using that data
\end{itemize}

\center{\textbf{Models}}

\begin{itemize}
    \item Supervised Learning
    \begin{itemize}
        \item Decision Trees
        \item Bayesian Networks
    \end{itemize}

    \item Unsupervised Learning
    \begin{itemize}
        \item K-Means
        \item Affinity Propagation
    \end{itemize}
\end{itemize}
\end{frame}

% =============================================================================
\begin{frame}
\frametitle{Background}

\center{\textbf{Educational Data Mining}}

\begin{itemize}
    \item Educational Data Mining consists in applying Data Mining methods to educational data
    \item Successful experiments have been made in which student grades are predicted from their usage of LMS systems
    \item The used models are often Decision Trees, Bayesian Networks, Neural Networks, among others
\end{itemize}

\end{frame}