### 2017-07-05 presentation/statistics.tex

\section{Statistics on the Data}

\begin{frame}
\frametitle{Statistics on the Data - Courses}

\begin{columns}[T] % align columns
\begin{column}{.48\textwidth}
    \begin{itemize}
        \item Courses in Numbers
    \end{itemize}

    \begin{tabular}{| l | r |}
        \hline
        Number of Courses     & 77 \\ \hline
        Number of Departments & 9  \\ \hline
        Number of Degrees     & 11 \\ \hline
    \end{tabular}

    \vspace{1cm}

    \begin{itemize}
        \item 28 out of the 77 courses are from Computer Engineering
        \item Most degrees have courses in both semesters
    \end{itemize}
\end{column}%
\hfill%
\begin{column}{.48\textwidth}
    \begin{itemize}
        \item Number of courses group by credits
    \end{itemize}

    \begin{tabular}{| l | r |}
        \hline
        Credits & Number of Courses \\ \hline
        1       & 3                 \\ \hline
        2       & 3                 \\ \hline
        3       & 3                 \\ \hline
        5       & 6                 \\ \hline
        6       & 46                \\ \hline
        7.5     & 12                \\ \hline
        12      & 2                 \\ \hline
        15      & 2                 \\ \hline
    \end{tabular}
\end{column}%
\end{columns}
\end{frame}

\begin{frame}
\frametitle{Statistics on the Data - Courses}

\centering
\includegraphics[width=9cm]{../src/stat_007_results/plot}

\begin{itemize}
    \item {\color{gray} Number of Students per Course} and {\color{blue} Number of Approved Students per Course}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Statistics on the Data - Moodle Users}

\begin{columns}[T] % align columns
\begin{column}{.48\textwidth}
    \begin{itemize}
        \item Moodle Users Counts
    \end{itemize}

    \begin{tabular}{| l | r |}
        \hline
        Number of Students   & 145 \\ \hline
        Number of Professors & 87  \\ \hline
        Total                & 232 \\ \hline
    \end{tabular}
\end{column}%
\hfill%
\begin{column}{.48\textwidth}
    \begin{itemize}
        \item Some similar studies have numbers of students between 800 and 900
        \item But they also have lower numbers of courses such as 7
    \end{itemize}
\end{column}%
\end{columns}
\end{frame}

\begin{frame}
\frametitle{Statistics on the Data - Moodle Users}

\center{\textbf{CRUD of Most Active Courses}}

\begin{columns}[T] % align columns
\begin{column}{.48\textwidth}
    \includegraphics[width=4.7cm]{../src/stat_004_results/fig_most_0_CTE-UMCED-2_all}

    \includegraphics[width=4.7cm]{../src/stat_004_results/fig_middle_0_MAT7514_1_all}
\end{column}%
\hfill%
\begin{column}{.48\textwidth}
    \includegraphics[width=4.7cm]{../src/stat_004_results/fig_most_1_eDocOn-T1_all}

    \includegraphics[width=4.7cm]{../src/stat_004_results/fig_middle_1_GES11656_all}
\end{column}%
\end{columns}
\end{frame}

\begin{frame}
\frametitle{Statistics on the Data - Moodle Users}

\center{\textbf{CRUD of Median Active Courses}}

\begin{columns}[T] % align columns
\begin{column}{.48\textwidth}
    \includegraphics[width=4.7cm]{../src/stat_004_results/fig_most_0_CTE-UMCED-2_cud}

    \includegraphics[width=4.7cm]{../src/stat_004_results/fig_middle_0_MAT7514_1_cud}
\end{column}%
\hfill%
\begin{column}{.48\textwidth}
    \includegraphics[width=4.7cm]{../src/stat_004_results/fig_most_1_eDocOn-T1_cud}

    \includegraphics[width=4.7cm]{../src/stat_004_results/fig_middle_1_GES11656_cud}
\end{column}%
\end{columns}
\end{frame}

\begin{frame}
\frametitle{Statistics on the Data - Moodle Users}
\centering

\center{\textbf{Total Activities per Week vs Mean Activities per Week}}

\includegraphics[width=9cm]{../src/stat_005_results/plot}
\end{frame}

\begin{frame}
\frametitle{Statistics on the Data - Moodle Users}
\centering

\center{\textbf{Total Number of Active Courses per Week (Bars) vs Total Active Courses (Line)}}

\includegraphics[width=9cm]{../src/stat_006_results/plot}
\end{frame}

\begin{frame}
\frametitle{Statistics on the Data - Results}

{

\center

\begin{tabular}{| l | r |}
    \hline
    Number of results           & 716 \\ \hline
    Number of non-empty results & 659  \\ \hline
    Number of approved students & 334 \\ \hline
\end{tabular}

}

\end{frame}